### 2015-09-15 ack.tex

%% Acknowledgments
\section{Acknowledgements}

%I would like to thank everybody who made it possible for me to write
%this paper. 
I would like to thank my mentor, 
{Norman Ramsey}
%\footnote{Norman Ramsey is a professor at Harvard University (Electrical
%Engineering and Computer Science).}
for offering me such a fascinating project and for giving me a hand when
all the theorems and proofs seemed to be incorrect. I also appreciate the
constructive criticism of my tutor, {Jenny Sendova} 
who was so kind as to read
through the drafts several times and tirelessly untangle my absolutely
confusing complex sentences. I thank {LeeAnn Tzeng} for her remarks
on the paper which helped me a lot to make it more understandable.

Also I would like to thank all sponsors and members of staff of the
{Research Science Institute} for making the research possible.