### 2015-09-15 conclusion.tex

%% Conclusion
\section{Conclusion}

We built an algebra of commands on filesystems. Then we proved that the
algebra is sound and complete, that is, commands which are algebraically
equivalent commands are also equivalent on real filesystems, and vice
versa.

We defined conflicts with the help of this algebra, then we defined
algorithms for update--detection and reconciliation. Finally, we
implemented these algorithms and tested it on numerous cases (see
Appendix \ref{app:impl}). The results were always in accordance with what
we expected from the philosophy ``propagate every command possible on
every replica''.  Because of that, using algebraical methods in file
synchronization turned out to be an effective solution.

We also gained some information about the provability of such theorems and
algebras. The methods we used in the proofs are likely be able to be used
in proofs for other algebras as well, so that we would be able to create a
more general theory of synchronization.