### 2015-09-15 intro.tex

%% Introduction
\section{Introduction}

If we have multiple copies (called \emph{replicas}) of a filesystem or a
part of a filesystem (for example on a laptop and on a desktop
computer), it is often the case that changes are made to each
replica independently and as a result they do not contain the same
information. In that case, a \emph{file synchronizer} is used to make them
similar (consistent) again, without losing any information.

The goal of a file synchronizer is to \emph{detect conflicting updates}
and \emph{propagate non--conflicting updates on each replica.} Due to the
various cases of file updates and the various behaviors of filesystem
commands it is difficult to give a complete and correct definition of its
behavior.

It can be helpful in solving this problem if we have a
formalized mathematical model for describing file synchronization.

We investigated algebraic models of synchronization in a specific case to
see the properties of such a system and the method of proving theorems in
it which will hopefully aid us to create a general theory of
synchronization based on algebraic structure in the future. This specific
case was the synchronization of filesystems.

In the paper we develop an algebraic representation of filesystem
commands. Then we show that this model is both complete and sound,
and after that we define the expected behavior of a file synchronizer and
create its specification with the help of our algebra.

This method will
hopefully simplify the specification as well as the implementation of the
synchronizer, and it might make it possible to extend the synchronizer to
other types of datasystems, such as mail folders or databases.

\subsection{The main idea of synchronization}

In general, two phases of a synchronizer's task can be separated: the
\emph{update detection}, when the program recognizes where updates have
been made since the last synchronization and \emph{reconciliation}, when
it combines updates to yield a new, synchronized version of each replica.
If there are no conflicting updates, then at the end of the
synchronization the replicas will be the same. Otherwise a synchronizer
may leave them unchanged at paths where conflicts occurred and warn the
user about them.

In this paper,
we focus on the commands that were applied to these replicas while they
were modified independently. The main aim of a synchronizer is to perform
\emph{all} commands on \emph{all} replicas.

%\remark{Explanation why we DO need O?}
Let us introduce some notation. We have the original system, \(O\); 
and the present replicas \(R_1, R_2, \ldots R_n\).
Somehow we know the sequences of commands which were applied to
\(O\) in each replica (for example from the update detector). They are
\(S_1, S_2, \ldots S_n\). The synchronizer algorithm will provide us
commands on each replica (\(S_1^*, S_2^*, \cdots S_n^*\)) which lead each
replica to a common state \(O'\) if possible.
%, that is, 
%\(S_1^*R_1=S_2^*R_2=\cdots=S_n^*R_n=O'\).

In order to determine each sequence \(S_i^*\), we take all commands
applied
to the systems (\(S_1\cup S_2\cup \cdots\cup S_n\)) then omit commands
in \(S_i\) which were already applied to replica \(R_i\).

But that way we only gain a \emph{set} of commands without order. We
must order it somehow to be able to apply the commands to the
system. But it is possible that some orders cause errors in a filesystem
(for example, trying to remove a directory before removing the files in
it). It is also possible that not all error--free orders have the same
effect on the filesystem. For example, modifying the contents of a file to
``xxx'' and modifying them to ``yyy'' are not commutable commands: they
leave the system in different states if applied in different order to it.
In this case the synchronizer cannot synchronize the systems fully, since
``last modifying wins'' is not a good solution in every case.  
Therefore the aim of the synchronizer is to find commands for which 
\emph{all error--free orders have
the same effect} and apply them to the system.

The reason we may have two differently ordered sequences of commands which
do not lead a system to the same state is that we have incommutable pairs
of commands in them. This is because if all pairs of commands
commuted (that is, pair of commands \(C_1; C_2\) had the same effect on
a system as \(C_2; C_1\)) then clearly one of the sequences could be
changed to the other one by commuting commands
without any change in the effect of the sequence on the system.

Because of this, we focused on the commutability of pairs of commands.  
We defined this property with the help of an algebraic 
proof system on commands.% for commutivity. %%%

%\subsection{Synchronizing filesystems}
%This specific case was the synchronization of a filesystem.