### 2015-11-22 2000/ack.tex

%% Acknowledgments
\section{Acknowledgements}

I would like to thank my mentor, 
{Norman Ramsey}
%\footnote{Norman Ramsey is a professor at Harvard University (Electrical
%Engineering and Computer Science).}
for allowing me to work on such a fascinating project,
and for always being ready to discuss and help me overcome
whatever difficulties I ran into during my work.
I also appreciate the
constructive criticism of my tutor, {Jenny Sendova} 
who kindly read my drafts
several times tirelessly untangling my 
sentences fraugh with overcomplicated grammar.
I thank {LeeAnn Tzeng} for her remarks
on the paper which helped to make it more accessible.

Also I would like to thank all sponsors and members of staff of the
{Research Science Institute} for making the research possible.