### 2015-12-21 2015/references.tex

\bibitem{NREC} 
Norman Ramsey and Elod Csirmaz,
``An Algebraic Approach to File Synchronization,''
in \emph{Proceedings of the Joint 8\textsuperscript{th} European Software Engineering Conference 
and 9\textsuperscript{th} ACM SIGSOFT Symposium on the Foundations of Software Engineering,}
New York: ACM, 2001, pp. 175--185.

\bibitem{CBNR} 
Constantin Berzan and Norman Ramsey,
``Summer Scholars Technical Report,''
Medford, MA: Tufts University, 2010.
Accessed 29 November 2015.
http://thirld.com/files/summerscholars\_techreport.pdf.

% 5
\bibitem{TTPDSH}
% D. B. Terry and M. M. Theimer and Karin Petersen and A. J. Demers and M. J. Spreitzer and C. H. Hauser
D. B. Terry et al.,
``Managing Update Conflicts in Bayou, a Weakly Connected Replicated Storage System,''
in \emph{Proceedings of the Fifteenth ACM Symposium on Operating Systems Principles,}
New York, NY: ACM, 1995, pp. 172--182.

% 4
\bibitem{BP}
S. Balasubramaniam and Benjamin C. Pierce,
``What is a File Synchronizer?''
in \emph{Proceedings of the 4\textsuperscript{th} Annual 
ACM/IEEE International Conference on Mobile Computing and Networking,}
New York: ACM, 1998, pp. 98--108.

% 3
\bibitem{KRSD}
Anne-Marie Kermarrec and Antony Rowstron and Marc Shapiro and Peter Druschel,
``The IceCube Approach to the Reconciliation of Divergent Replicas,''
in \emph{Proceedings of the Twentieth Annual ACM Symposium on Principles of Distributed Computing,}
New York: ACM, 2001, pp. 210--218.