### 2015-12-22 2015/paper.tex

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{centernot}
\usepackage{tikz}
\usepackage{cleveref}

\input{macros}

\title{Complete Algebraic File Synchronization}
\author{Elod Pal Csirmaz}

\begin{document}
\maketitle
\begin{abstract}
Abstract goes here
\end{abstract}

\input{intro}

\section{Definitions}\label{sec_def}

\input{def-filesystem}

\input{def-commands}

\input{axioms}

\section{Update Detection}\label{sec_update}

\input{upd-def}

\input{ordering}

\input{upd-correct}

\section{Reconciliation}\label{sec_rec}

\input{rec-intro}

\input{works-inverse}

\input{rec-sound}

\input{rec-complete}

\input{comparisons}

\section{Further Research}

TODO: next steps. multiple replicas

\input{algebra}

\section{Conclusions}

% TODO
Conclusions go here

\section{Acknowledgments}

I thank Laszlo Csirmaz for his invaluable input on previous drafts of this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.

\begin{thebibliography}{99}

\input{references}

\end{thebibliography}

\end{document}