### 2016-01-03 2015/abstract.tex
New title, keywords, MSC, ACM classes

With distributed computing and mobile applications,
synchronizing divergent replicas of data is a more and more common problem.
We investigate an operation-based approach to filesystem synchronization
and further our theoretical understanding of this problem
by presenting rigorous algebraic proofs that the algorithms described for
update detection and reconciliation return working solutions
that cannot be improved on.

To do this, we present our filesystem model 
and a novel, symmetric set of filesystem commands
with higher information content,
which allow us to simplify our definition of conflicting updates
compared to our previous approach,
and to ensure that we propagate all updates that do not cause errors.
We also present a number of generally useful properties of sequences of commands
including introducing the concept of inverse sequences and conditional operation.

These results are essential in
determining the outcome of an
application implementing an algebraic approach to synchronization
and in ensuring that it is error-free,
and as a step towards building a more complete algebra
to describe updates to filesystems and other storage protocols.