### 2016-01-03 2015/paper.tex

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{centernot}
\usepackage{tikz}
\usepackage[capitalise]{cleveref}

\input{macros}

% fitness, validity, integrity, well-founded
% foundation, groundwork
\title{New Foundations and Fitness of Algebraic File Synchronization}
\author{Elod Pal Csirmaz\\
\texttt{\rot{\rot{maz.}{csir}{ep}com}{@}{elod}}}

\begin{document}
\maketitle
\begin{abstract}

\input{abstract}

\bigskip\noindent
{\bf Keywords:} % (4-6)
file synchronization,
algebraic approach,
distributed systems,
reconciliation,
data replication
% conflict detection,
% operation-based 
% conflict resolution
% computer algebra

\bigskip\noindent
{\bf MSC classes:} % http://www.ams.org/mathscinet/msc/msc2010.html
08A70, % Applications of universal algebra in computer science
68M07, % Mathematical problems of computer architecture
68M14, % Distributed systems
68P05, % Data structures
68P20 % Information storage and retrieval
% 68N25         Operating systems

\bigskip\noindent
{\bf ACM classes:} % http://www.acm.org/about/class/ccs98-html
% (D) Software (D.4) Operating systems
D.4.3, % File systems management
% (E) Data
E.5, % Files
% (F) Theory of computation (F.2) ANALYSIS OF ALGORITHMS AND PROBLEM COMPLEXITY
F.2.2, % Nonnumerical Algorithms and Problems
% (G) Mathematics of Computing
G.2 % DISCRETE MATHEMATICS
% (H) Information Systems (H.3) INFORMATION STORAGE AND RETRIEVAL
% H.3.4 % Systems and Software

\end{abstract}

% section
\input{intro}

\section{Definitions}\label{sec_def}

\input{def-filesystem}

\input{def-commands}

% section
\input{axioms}

\section{Update Detection}\label{sec_update}

\input{upd-def}

\input{ordering}

\input{upd-correct}

\section{Reconciliation}\label{sec_rec}

\input{rec-intro}

\input{works-inverse}

\input{rec-sound}

\input{rec-complete}

% section
\input{comparisons}

% section
\input{algebra}

% section
\input{conclusion}

\section{Acknowledgments}

I thank Laszlo Csirmaz for his invaluable input on previous drafts of this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.

\begin{thebibliography}{99}

\input{references}

\end{thebibliography}

\end{document}