### 2016-01-04 i/lawproof.tex
Renamed directories

%% Example proof for law
\section{Soundness proofs for individual laws}
\label{app:soundness}

%{\small{
In this section we provide examples how algebraic laws can be proved.
The method we use is to consider as many cases as necessary to be able to
predict the result of each command in the law. For example, if the law
contains paths \(\pi\) and \(\pisub\) then we'll consider 
22 cases; see below.
\remark{...}
(Note that we listed all cases considering how \(\pisub\) can relate to
\(\pi\) and if they have children or a parent. Also, we considered the
case when \(F(\pi)\) has
only one child and it is \(F(\pisub)\) since in that case, modifying
\(F(\pisub)\) can make \(F(\pi)\) childless.)

If the law contains an \(edit\) command, we also distinguish
between files and directories. If we determine the result of the sequence
of commands of the left side in each case and this is the same as that of
the right side, then the law holds.

Now we provide an example proof on law 18. We will only investigate the
left side of the law, since the right side always gives the broken
filesystem.

\begin{forrsi*}
\begin{singlesp}
\end{forrsi*}
\begin{multicols}{3}
{\tiny{
\input{cases}
}}
\end{multicols}
\begin{forrsi*}   
\end{singlesp}
\end{forrsi*}
%}}