### 2016-01-05 p1/conclusion.tex
Renamed directories

%% Conclusion
\section{Conclusion}

We have built an algebra of commands on filesystems. We have proved that the
algebra is sound and complete, that is, sequences of commands which are algebraically
equivalent are also equivalent on real filesystems, and vice
versa.

We have defined conflicts with the help of this algebra, and defined
algorithms for update detection and reconciliation. We have also
implemented these algorithms and tested them on numerous cases (see
Appendix \ref{app:impl}). The results have always been in accordance with what
we expected from the philosophy ``propagate every command possible on
every replica''.  Using algebraical methods in file
synchronization has turned out to be an effective solution.

We have also gained some information about the provability of such theorems and
algebras. The methods we have used in the proofs are likely be able to be used
in proofs for other algebras as well, thus they may form the basis for
a more general theory of synchronization.