### 2016-01-05 p2/abstract.tex

With distributed computing and mobile applications,
synchronizing divergening replicas of data structures is a more and more common problem.
We use algebraic methods to reason about filesystem operations, 
and introduce a simplified definition of conflicting updates to filesystem trees.
We also define algorithms for update detection and reconciliation
and present rigorous proofs that they not only work as intended,
but also cannot be improved on.

To achieve this, we introduce a novel, symmetric set of filesystem commands
with higher information content,
which removes edge cases
and increases the predictive powers of our algebraic model.
We also present a number of generally useful classes and properties
of sequences of commands.

These results are often intuitive,
but providing exact proofs for them is far from trivial.
They contribute to our understanding of this special type of algebraic model,
and toward building a more complete algebra
of filesystems and other storage protocols.
They also form a theoretical basis for
specifying
and guaranteeing the error-free operation
of applications that implement an algebraic approach to synchronization.