### 2017-10-15 p2/unused.tex
Inverses are no longer needed

\section{UNUSED RESULTS}

% Type equivalence classes
% ------------------------

\noindent
For the proof of the correctness of the reconciliation algorithm,
we will also need a way of moving parts of sequences between the two
sides of the $\worksmeqsign$ relation.
We can do this based on the observation that commands are bijections
if we only distinguish between types in our objects.
To formalise this, we extend type equality ($\typeeq$)
to filesystems, commands and sequences of commands.
\begin{mydef}[Equivalence by type, extended: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems in $\setfs$, $\FS\typeeq\GS$ iff
for all $n\in\setn$, $\FS(n) \typeeq \GS(n)$.
\item For two commands $\cxynv\typeeq\czwmv$ iff 
$x\typeeq z$ and $y\typeeq w$
(that is, $[x]=[z]$ and $[y]=[w]$)
and $n=m$.
\item To sequences of commands, the relation $\typeeq$ extends as usual: $\emptyseq\typeeq\emptyseq$, and
if $A'=\alpha\cc A$ and $B'=\beta\cc B$, then $A'\typeeq B'$ iff $\alpha\typeeq\beta$ and $A\typeeq B$.
\end{itemize}
\end{mydef}

We select representatives from each $\typeeq$-equivalence class of values:
for $[\empt]$ and $[\vald]$ this must be $\empt$ and $\vald$; and
for $[\valf]$ we select an arbitrary file value $\fil\in[\valf]$.
We then use these representative values to
introduce the \emph{canonical mapping} $\T{}$ that preserves the types of values,
but maps all objects into a subsystem where there is a one-to-one
correspondence between types and values.

\begin{mydef}[canonical mapping, $\T{}$]\label{def_typemapping}
We use $\T{}$ and write $\T{a}$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\T{\vald}=\vald$, $\T{\valf}=\fil$ and $\T{\empt}=\empt$.
\item For filesystems, $\T{\FS}$ is defined by $\T{\FS}(n) = \T{\FS(n)}$.
\item For commands, $\T{\cbrk}=\cbrk$ and $\T{\cxynv} = \cTxynv$.
\item To sequences of commands, the mapping extends as usual: $\T{\emptyseq}=\emptyseq$, and if $S'=\alpha\cc S$, then $\T{S'}=\T{\alpha}\cc \TS$.
\item Finally we extend $\T{}$ to sets of sequences: $\T{\sqs{A}} = \{\TS \whr S\in\sqs{A}\}$.
\end{itemize}
\end{mydef}

We refer to any object in the image of $\T{}$ as \emph{canonical}.

By definiton,
for any filesystem $\FS$ and command $\alpha$, 
$\T{\alpha}(\T{\FS}) = \T{(\alpha\aFS)}$.

% TODO ???
Also, for all objects where it is defined, $\T{a}\typeeq a$.
As canonical filesystems, commands, sequences, etc.
are subsets of their non-canonical counterparts, 
all propositions proven so far apply to them as well.

As the types of values determine whether a command breaks a filesystem,
we know $\Dom{A}=\Dom{\TA}$ and therefore $\Dom{\sqs{A}}=\Dom{\T{\sqs{A}}}$.
From this it follows that
% TODO what's the set of all filesystems?
% TODO:
\begin{mycor}\label{repr_works_is_same}
\begin{gather*}
\worksmnil{\sqs{A}} \Longleftrightarrow \worksmnil{\T{\sqs{A}}} \\
\worksm{\sqs{A}}{\sqs{B}} \Longleftrightarrow \worksm{\T{\sqs{A}}}{\T{\sqs{B}}}.
\end{gather*}
\end{mycor}

% Inverse commands
% ----------------

We are now ready to define the inverse of canonical commands $\T{\alpha}$
which allows us to move parts of sequences between the
two sides of $\worksmeqsign$.
This is based on the following observation.

\begin{mylem}\label{repr_comm_inject}
Canonical commands $\T{\alpha}$ are partial bijections over canonical filesystems
($\T{\FS}$, $\T{\GS}$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that canonical functions are injective except for cases
where they are not defined:
\begin{align*}
\forall\FS,\GS,\alpha:& \\
& \T{\FS}\neq\T{\GS} \wedge \T{\alpha}(\T{\FS})\neq\fsbroken \\
& \Rightarrow \T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS}).
\end{align*}
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\T{\alpha} = \cTxynv$.
If $\T{\FS}$ and $\T{\GS}$ have different values (types) at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\T{\FS}$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\T{\FS})\neq\fsbroken$, we know $[\FS(n)]=[x]$.
We also know that $\T{\GS}(n)\neq\T{\FS}(n)$, and therefore
$[\GS(n)]\neq[x]$.
This implies $\cTxynv(\T{\GS})=\fsbroken$,
and again we have $\T{\alpha}(\T{\FS}) \neq \T{\alpha}(\T{\GS})$.
\end{proof}

% See
% https://en.wikipedia.org/wiki/Inverse_semigroup#Origins
% https://en.wikipedia.org/wiki/Symmetric_inverse_semigroup
% https://en.wikipedia.org/wiki/Partial_function

\begin{mydef}[Inverse commands and sequences]
We write $\T{\alpha}^{-1}$ for
the inverse of the canonical command (partial bijection) $\T{\alpha}$,
which is uniquely defined by \cref{repr_comm_inject}.

We write $\TS^{-1}$ for the inverse of the sequence $\TS$, which consists of the inverses of the commands in $\TS$ in reverse order.
\end{mydef}
By definition, for any canonical filesystem $\T{\FS}$ outside the image of 
$\T{\alpha}$, the inverse $\T{\alpha}^{-1}$ is not defined.

% Some further observations:
% For canonical commands \[ \cTxynv^{-1} = \cTyxnv. \]
% $\forall S: ({\TS}^{-1})^{-1} = \TS$.

For brevity, in the following lemma we use $B$, $B'$, $\sqs{A}$ and $\sqs{C}$ to denote
canonical sequences and sets.

\newcommand{\ia}{\sqs{A}}
\newcommand{\ic}{\sqs{C}}
\newcommand{\ibi}{B^{-1}}

\begin{mylem}\label{r_invmove}
A common initial segment of a set of canonical sequences can be moved 
to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\ia}{B\cc\ic}
\ \Longleftrightarrow \ \worksm{\ibi\cc\ia}{\ic} 
\ \wedge \  \worksm{\ia}{B}
\end{gather*}
\end{mylem}
\begin{proof}
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

The proof is illustrated by \cref{fig_invmove}.

\begin{figure}[htb]
\input{fig-inverse-works}
\caption{Proof of \cref{r_invmove}}\label{fig_invmove}
\end{figure}

For the left-to-right part of the proposition,
from $\worksm{\ia}{B\cc\ic}$ we know $\Dom{\ia} \subseteq \Dom{B\cc\ic}$.
As the (canonical) $B$ is a bijection between its domain and image,
we can use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc\ia} \subseteq \Dom{\ibi\cc B\cc\ic}$,
but the latter is a subset of $\Dom{\ic}$,
and so $\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$,
which means that $\worksm{\ibi\cc\ia}{\ic}$.
It is easy to see that $\worksm{\ia}{B\cc\ic} \Rightarrow \worksm{\ia}{B}$.

For the right-to-left part,
from $\worksm{\ia}{B}$ we know $\Dom{\ia} \subseteq \Dom{B}$.
Therefore we can project the whole of $\Dom{\ia}$ using the bijective canonical sequence $B$
to get $\Dom{\ibi\cc\ia}$, and as $\worksm{\ibi\cc\ia}{\ic}$
we know
$\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$.
We use $\ibi$ to project both sides of this statement to get
$\Dom{B\cc\ibi\cc\ia} \subseteq \Dom{B\cc\ic}$.
But we also know that $\Dom{B\cc\ibi\cc\ia} = \Dom{\ia}$
as $\Dom{\ia} \subseteq \Dom{B}$,
and so $\Dom{\ia} \subseteq \Dom{B\cc\ic}$ which means
$\worksm{\ia}{B\cc\ic}$.
\end{proof}

We use this result to prove the following:

\begin{mylem}\label{indep_prefix_combine}
The combination of two sequences with a common head and independent tails 
is defined wherever the two sequences are defined:
% Use A\cc\{B,C\}
\[ B\indep C \Rightarrow \worksmbx{A\cc B, A\cc C}{A\cc B\cc C} \]
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnil{A}$, so we assume that that is the case.
From $B\indep C$ and \cref{combine_independent_sequences} we know
$\worksmbx{B, C}{B\cc C}$.
Based on \cref{repr_works_is_same} this is equivalent to
$\worksmbx{\TB, \TC}{\TB\cc \TC}$.
Since $\TA^{-1}\cc \TA \equiv \emptyseq$, this can also be written as
\[\worksm{\TA^{-1}\cc\{\TA\cc \TB, \TA\cc \TC\}}{\TB\cc \TC}.\]
From \cref{worksextpostfix} we also know that
$\worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA}$.
Combining the last two statements using \cref{r_invmove}
we get
\[ \worksmbx{\TA\cc \TB, \TA\cc \TC}{\TA\cc \TB\cc \TC}, \]
which with \cref{repr_works_is_same} proves the
\namecref{indep_prefix_combine}.
\end{proof}