### 2017-10-19 p2/rec-complete.tex
Completed the completeness theorem

\subsection{Reconciliation is Maximal}

% Reconciliation is maximal
% -------------------------
% This is where we're using |D|=1

%% TODO general properties of commands ==> behavior on specific filesystems

The reconciliation algorithm defined above is also maximal, that is,
it is not possible to apply any further commands from $\ambnp$ to $\FS_B$.
To show this, we are going to prove that any sequence $S$ formed from commands
in $\ambnp$ that is not independent of $\bmanp$ necessarily breaks $B\aFS=\FS_B$
or introduces a conflicting update.

\begin{myth}\mlabel{rec_is_complete}
If $A$ and $B$ are simple sequences,
$S'\in\orderset{\ambnp}$,
and $S$ is a subsequence of $S'$ that
contains a command $\cxynv$ for which $\cxynv\nindep\bmanp$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ are defined on
(that is, $S$ breaks all possible $\FS_B$ replicas),
or $S$ changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but to values of the same type), which 
our algorithm must treat as a conflict to be resolved by
by the user or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Let $T$ be an arbitrary sequence in $\orderset{\bmanp}$.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $T$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep T$.

Let $\czwmv$ be the last command in $T$ that is not independent of $\cxynv$,
and split $T$ into $T_0\cc\czwmv\cc T_1$, where therefore $T_1\indep\cxynv$.

From \cref{can_move_intersection} we know that
$\acb\cc\bma \in \orderset{B}$, and so
$B\cc S \equiv \acb\cc\bma\cc S \equiv \acb\cc T\cc S$.
We also know that commands in $S_0$ and $T$ commute, and so this is equivalent to
$\acb\cc S_0\cc T\cc\cxynv\cc S_1$.
Expanding $T$ and swapping $T_1$ and $\cxynv$ we get
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv\cc T_1\cc S_1. \]

First, we prove that this sequence breaks all filesystems
where $A$ and $B$ are defined unless $n=m$.
Let us therefore suppose $n\neq m$ and, to use an inverse proof,
that the sequence is defined on $\FS$ 
where $A\aFS\neq\fsbroken$ and $B\aFS\neq\fsbroken$.
If so, its initial segment,
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv, \]
must also be defined on $\FS$.

We know $\cxynv\nindep\czwmv$, and so from \cref{incomparable_is_independent}, $n\nunrel m$.
As the sequence is defined on $\FS$,
because of \cref{simple_distant_pairs}
$\czwmv\cc\cxynv$ can only be a construction or destruction pair.

Also, because $B$ is simple and contains $\czwmv$,
we know there are no commands on $m$ 
in either $\acb$ or $T_0$, which are both formed from other commands in $B$.
Moreover, we know there are no commands on $m$ in $S_0$ because $S_0\indep T$.
We therefore know that $\FS(m)$ must be of type $[z]$.

If $m=\parentf{n}$ then $\czwmv\cc\cxynv$ is a construction pair, and we know
$[\FS(m)]=[z]\neq[\vald]$ and $y\neq\empt$ as none of these commands is an assertion command.
If so, $\fsreplacement{\FS}{n}{y}$ 
(that is, applying $\cxynv$ to $\FS$ without modifying $\FS(m)$ first) would violate the tree property.
The same is true when $\parentf{m}=n$, where we have a destruction pair,
and $[z]\neq\empt$ and $[y]\neq[\vald]$.

Since $\cxynv\in S$ and therefore $\cxynv\in A$, and $A\aFS\neq\fsbroken$,
the above means that $A$ must also contain a command on $m$ before $\cxynv$.
From \cref{simple_distant_pairs} we then know that $\cqrmv\cc\cxynv$
is also a construction or destruction pair, and it is the same type as
$\czwmv\cc\cxynv$.

Because the type of the pair determines the output type of the first command,
this means that $[r]=[w]$, and they can either be $[\empt]$ or $[\vald]$.
As we assumed that $|\setd|=1$, we therefore know that $r=w$.
From \cref{worksinputmatch} we also know that $[q]=[z]$.

Therefore $\cqrmv = \czwmv$, which is a contradiction
because $\cqrmv$ was selected from $A$, and
$\czwmv$ was selected from $T$, that is, from $\bmanp$.

The only possibility is therefore $n=m$.
As $A$ and $B$ are simple, we then know that
$\cxynv$ and $\czwmv$ are the first commands on $n$,
and so (again from \cref{worksinputmatch}) $[z]=[x]$.
We also know $w\neq y$ as otherwise the two commands would be equal and
would be in $\acbnp$.
This, however, means that $\cxynv$, from $A$, overrides a change introduced 
by $\czwnv$, from $B$, which must be treated as a conflict.
\end{proof}

From the proof it is apparent that this result depends on $|\setd|=1$.