### 2018-07-08 p2/multidir.tex

\section{Extending the Synchronizer for Directory Metadata}\mlabel{sec_multidir}

As suggested above, one of the central assumptions of our model is that
there is only one directory value, that is, directories are not differentiated
by the meta-information they contain.
In this section we describe reasons why this may not limit the applicability
of even the current model, and we describe pre- and post-processing steps
to overcome this limitation.

% TODO