### 2018-07-08 p2/paper.tex

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{centernot}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage[capitalise,poorman,nameinlink]{cleveref}

\input{macros}
\input{axiom-def}

\title{Algebraic File Synchronization: Adequacy and Completeness}
\author{Elod Pal Csirmaz\\
\texttt{\rot{\rot{maz.}{csir}{ep}com}{@}{elod}}}
\date{}

\begin{document}
\maketitle
\begin{abstract}

\input{abstract}

\myskip
{\bf Keywords:} % (4-6)
file synchronization,
algebraic approach,
distributed systems,
reconciliation,
data replication
% conflict detection,
% conflict resolution
% computer algebra

\myskip
{\bf MSC classes:} % http://www.ams.org/mathscinet/msc/msc2010.html
08A70, % Applications of universal algebra in computer science
68M07, % Mathematical problems of computer architecture
68M14, % Distributed systems
68P05, % Data structures
68P20 % Information storage and retrieval
% 68N25         Operating systems

\myskip
{\bf ACM classes:} % http://www.acm.org/about/class/ccs98-html
% (D) Software (D.4) Operating systems
D.4.3, % File systems management
% (E) Data
E.5, % Files
% (F) Theory of computation (F.2) ANALYSIS OF ALGORITHMS AND PROBLEM COMPLEXITY
F.2.2, % Nonnumerical Algorithms and Problems
% (G) Mathematics of Computing
G.2 % DISCRETE MATHEMATICS
% (H) Information Systems (H.3) INFORMATION STORAGE AND RETRIEVAL
% H.3.4 % Systems and Software

\end{abstract}

% section
\input{intro}

\section{Definitions}\mlabel{sec_def}

\input{def-filesystem}

\input{def-commands}

% section
\input{axioms}

\section{Update Detection}\mlabel{sec_update}

\input{upd-def}

\input{ordering}

\input{upd-correct}

\section{Reconciliation}\mlabel{sec_rec}

\input{rec-intro}

\input{works-inverse}

\input{rec-sound}

\input{rec-complete}

%section
\input{multidir}

% section
\input{algebra}

% section
\input{conclusion}

\section{Acknowledgments}

I thank L.~Csirmaz for his invaluable input on this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.
I am indebted to the anonymous reviewers of the previous version of this paper
for their helpful and detailed remarks
that helped improve the presentation of the paper
and make it accessible for a wider audience.

\begin{thebibliography}{99}

\input{references}

\end{thebibliography}

\appendix

% section
\input{comparisons}

\section{Proofs for Rules}\mlabel{axiom_proof}
\input{axiom-proof}

\end{document}