### 2018-07-09 p2/axiom-proof.tex
Simplified the proofs of rules

In this section we derive the \namecrefs{ax_separate_commute} listed in
\cref{rules_lemma} from our filesystem model.
In the proofs, we may say that a filesystem
is \emph{broken at $n$}
if after applying a command, it violates the tree property
because
$n$ is not a dictionary, but has at least one non-empty child.

% For the purposes of the proofs we extend the notion of filesystems
% to the full set of functions mapping $\setn$ to $\setv$, even if they
% do not have the tree property.
% This is so that we can describe the values they would acquire under certain nodes,
% and pinpoint where they may be violating the tree property.
% If they do, we continue to say that the filesystem is broken,
% and say that the filesystem  if 
% We continue to require that filesystems are not empty at finitely many nodes.
% 
% Accordingly, where needed, we also implicitly extend filesystem commands
% so that they would return a filesystem even if it violates the tree property.
% We continue to say that they break the filesystem in these cases,
% and commands continue to be undefined where their input types are incompatible
% with the input filesystem.
% We note that once a filesystem is broken, it stays broken, even if
% after a subsequent command it would no longer violate the tree property.

It follows from the definition of the tree property that
whether a filesystem is \emph{broken at a node $n$} is determined only
by the values at $n$ and the children of $n$.
% In other words,
% 
% \begin{myclm}[Locality of tree property violations]\mlabel{broken_local}
% If two filesystems over $\setn$ have the same values at a node $n$,
% and at all children of $n$ (if they exist),
% then either both filesystems are broken at $n$, or none of them are.
% \end{myclm}

Below we restate the \namecrefs{ax_separate_commute} defined in
\cref{rules_lemma}.

\begin{myaxproof}
% ax_separate_commute
\axaxseparatecommute
\end{myaxproof}

\begin{proof}
This proof is similar to the first case in the proof of \cref{combine_independent_commands}.
However, this result is needed for \cref{independent_details} on which that proof rests.

We use an inverse proof and assume $\cxynv\cc\czwmv \nequiv \czwmv\cc\cxynv$.
Since the values in the resulting filesystems are the same,
this is only possible if, for an initial non-broken filesystem $\FS$,
one side results in a broken filesystem, and the other side does not.
Without loss of generality, we can assume
$(\cxynv\cc\czwmv)\aFS$ is not broken, but $(\czwmv\cc\cxynv)\aFS$ is.
This means that either $\czwmv\aFS$ is already broken, or it is not, and applying $\cxynv$
breaks the filesystem.

In the former case 
$\czwmv\aFS$ must be broken at $m$ or $\parentf{m}$, since it only changed at $m$.
Let us look at the first case.
Since $n\indep m$, we know that the parent and children of $m$ have the same value
in $\FS$ as they do in $\cxynv\aFS$. However, this is a contradiction,
as applying $\czwmv$ to $\FS$ leads to a broken filesystem,
but applying it to $\cxynv\aFS$ (from the left side) does not.

Therefore $\czwmv\aFS$ must be broken at $\parentf{m}$.
By definition this means that its value at $\parentf{m}$ cannot be a directory.
Also, reasoning similar to the above shows that this is only possible if $\cxynv$ changes
the environment of $\parentf{m}$, which, since $n\indep m$, is only possible
if $n$ and $m$ are siblings.

Since then the value at $\parentf{n}=\parentf{m}$ is not changed by either command,
we know it cannot be a directory in $\FS$, either, and that therefore
$\FS$ is empty at all children of the parent node.
Since by assumption $\czwmv\aFS$ is broken, $[w]\neq[\empt]$ must hold.
However, this is a contradiction, as this would necessarily mean that
the left side, $\czwmv(\cxynv\aFS)$ must also be broken.

We can proceed in the same fashion if $\czwmv\aFS$ is not broken, but
$(\czwmv\cc\cxynv)\aFS$ is.
\end{proof}

\begin{myaxproof}
% \cref{ax_separate_nobreaks}
\axaxseparatenobreaks
\end{myaxproof}

\begin{proof}
It is easy to see that over any $\setn$ one can construct a filesystem that neither command breaks.
Since $n\indep m$, it is always possible to set all descendants of both $n$ and $m$ empty,
and have a directory at all ancestors of both $n$ and $m$. In such positions any values are permissible,
so neither command will break the filesystem.
\end{proof}

\begin{myaxproof}
% \cref{ax_same_breaks}
\axaxsamebreaks
\end{myaxproof}

\begin{proof}
This is trivial, as either $\cxynv$ breaks the filesystem, or, if it does not, then
we know $[\cxynv\aFS(n)]=[y]\neq [z]$, and therefore $\czwnv$ will break the filesystem.
\end{proof}

\begin{myaxproof}
% \cref{ax_same_emptyseq}
\axaxsameemptyseq
\end{myaxproof}

\begin{proof}
From the conditions we see that for every $\FS$,
$\FS$ and $(\cxynv\cc\czwnv)\aFS$ have the same values at every node,
while the latter can still be broken if $\cxynv\aFS$ is broken.
This is equivalent to saying that where the $\cxynv\cc\czwnv$ function is defined,
it is the identity function, that is, it is extended by $\emptyseq$.
\end{proof}

\begin{myaxproof}
% \cref{ax_same_singlec}
\axaxsamesinglec
\end{myaxproof}

\begin{proof}
It is easy to see that where both $\cxynv\cc \czwnv$ and $\cxwnv$ are defined
(they do not break the filesystem), they are equivalent.
What remains to show is that they break the same set of filesystems.
If $(\cxynv\cc \czwnv)\aFS$ is not broken, then we know that neither of
$\FS$, $\fsreplacement{\FS}{n}{y}$ or $\fsreplacement{\FS}{n}{w}$ is broken,
and therefore $\cxwnv$ does not break $\FS$.
Conversely, if $\cxwnv$ does not break $\FS$, then 
we know that $\FS$ and $\fsreplacement{\FS}{n}{w}$ are not broken.
Also, as
either $[x]\neq[\empt]$ or $[w]\neq[\empt]$,
there must be a directory at $\FS(\parentf{n})$,
and so $\fsreplacement{\FS}{n}{y}$ cannot be broken, either,
as any value is permissible at $n$.
This is equivalent to saying that $\cxynv\cc \czwnv$ does not break the filesystem.

We note that if $\cxwnv$ is an assertion command, then it extends $\cxynv\cc\czwnv$
instead of being equivalent to it.
\end{proof}

\begin{myaxproof}
% \cref{ax_directchild_breaks}
\axaxdirectchildbreaks
\end{myaxproof}

\begin{proof}
If neither of the commands is an assertion command, then the claim follows from
\cref{simple_distant_pairs} in the following way.
We use an inverse proof and assume that $\cxypnv\cc\czwnv$ is not a construction pair,
none of the commands is an assertion command, and $\wrksx{\cxypnv\cc\czwnv}$.
Therefore \cref{simple_distant_pairs} applies to this sequence, and the two commands
must form a construction or destruction pair. This is a contradiction as due to
the relationship between the nodes, they cannot form a destruction pair, and
they do not form a construction pair by assumption.

Otherwise, let $\FS_1$ be $\cxypnv\aFS$, and $\FS_2$ be $(\cxypnv\cc\czwnv)\aFS$.
If $\cxypnv=\cbba{\pn}$ (the only assertion command it can be given the conditions),
then either $\FS_1$ is broken, or
$\FS(\pn)=\FS_1(\pn)=\FS_2(\pn)=\empt$.
Since $\czwnv\neq\cbba{n}$, we know that either $\czwnv$ breaks $\FS_1$,
or, depending on $z$ and $w$, either $\FS_1(n)\neq\empt$ or $\FS_2(n)\neq\empt$.
In any case $\FS_2$ will be broken at $\pn$ which is empty but at one point has a non-empty child.

Finally, if $\czwnv=\cdda{n}$, we can proceed in a similar way,
as since $\cxypnv\neq\cdda{n}$, we know at one point the value at $\pn$ is not a directory.
\end{proof}

\begin{myaxproof}
% \cref{ax_directparent_breaks}
\axaxdirectparentbreaks
\end{myaxproof}

\begin{proof}
This proof follows the same logic as that of \cref{ax_directchild_breaks},
using destruction pairs instead of construction pairs.
\end{proof}

\begin{myaxproof}
% \cref{ax_distantrel_breaks}
\axaxdistantrelbreaks
\end{myaxproof}

\begin{proof}
We use an inverse proof and
suppose $\cxynnv$ and $\czwnv$ satisfy the conditions,
and there is an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czwnv\neq\cbba{n}$, then either $\FS(n)\neq\empt$ or $\czwnv\aFS(n)\neq\empt$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxynnv\neq\cdda{\nn}$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satisfy the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.
\end{proof}

\begin{myaxproof}
% \cref{ax_child_assert}
\axaxchildassert
\end{myaxproof}

\begin{proof}
Since $\cxynnv\neq\cdda{\nn}$, we know that
either $\cxynnv\aFS$ is broken, or, depending on $x$ and $y$,
either $\FS(\nn)\neq\vald$, or $\cxynnv\aFS(\nn)\neq\vald$.
In the latter two cases $\FS$ must be empty at all descendants of $\nn$
for it to satisfy the tree property.
The assertion command $\cbba{n}$ will therefore not break $\FS$.
\end{proof}

\begin{myaxproof}
% \cref{ax_parent_assert}
\axaxparentassert
\end{myaxproof}

\begin{proof}
This proof follows the same logic as that of \cref{ax_child_assert}.
\end{proof}

\begin{myaxproof}
% \cref{ax_assert}
\axaxassert
\end{myaxproof}

\begin{proof}
The two assertion commands $\cdda{n}$ and $\cbba{n}$ either
breaks a filesystem, or leaves it in the same state.
\end{proof}