### 2018-07-10 p2/conclusion.tex
Added conclusion

\section{Conclusions and Further Research}\mlabel{sec_conclusion}

In this paper we presented an algebraic model to file synchronization,
and defined an update detector and a reconciliation algorithm.
With its extended command set the model was carefully designed to take
advantage of an inherent symmetry of filesystem commands,
which was enhanced by removing meta-information from directories
as well as the \emph{move} command.
These assumptions enabled us to prove that
the update detector algorithm is correct (\cref{update_works}),
and that reconciliation is not only correct (\cref{reconciliation_correct})
but also yields maximal results (\cref{rec_is_complete}).

We introduced relations and concepts that have proven particularly
useful, including 
independent commands, % (\cref{def_indep}), 
simple sequences, % (\cref{def_min_simp}),
and relationships between domains of sets of command sequences ($\worksmeqsign$)
that allowed us to formalize the correctness of reconciliation.
Among our partial results were interesting properties of these concepts,
for example,
that all legal permutations of simple sequences are equivalent (\cref{simple_reorder_equiv}),
that command sequences can be transformed into simple sequences using syntactical rules (\cref{can_simplify}),
or that the combination of independent sequences is defined wherever the components are defined (\cref{combine_independent_sequences}).
We believe these results will prove useful outside their current usage as well.

Our results can be immediately applied to practical filesystem synchronizers,
which could benefit from e.g. recording changes using the extended command set,
and from the theoretical background that guarantees their correct operation.
Furthermore, with the encoding and decoding steps described in \cref{sec_multidir},
the reconciliation algorithm can be extended to handle directory meta-information
and the \emph{move} command,
which makes our model applicable to more use cases and filesystems.

\myskip
Apart from constructing a filesystem-free algebra,
there are many other ways in which further work can extend the current results.
An important extension would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed,
or show that it is impossible to satisfy these criteria.

A related problem is to extend the system and the proofs
to allow for cases where reconciliation cannot
complete fully
or if only a subset of the replicas are reconciled 
(e.g. due to network partitioning),
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystems.
Existing research can offer pointers as to how such cases can be modeled
in our algebraic system.
Parker et al. \cite{PPRS} and Cox and Josephson \cite{CJ}
describe version vectors (update histories) kept as metadata,
while Chong and Hamadi present distributed algorithms that allow incremental synchronization \cite{CH}.
Representing individual updates to files
in their modification histories (as described in \cite{CJ})
as separate commands could also enable an algebraic synchronizer to reconcile otherwise
conflicting updates and resolve partial reconciliations.

Future work could also investigate extending the 
model and algorithms to the $|\setd|>1$ case
so that directory metadata could be represented directly
as opposed to through encoding and decoding steps.

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of the update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and domains of sets of sequences ($\worksmeqsign$)
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.